---
title: "academic risk"
author: "Daniel Moscoe"
date: "10/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
loadObject <- function(fileName){
  load(fileName)
  get(ls()[ls() != "fileName"])
}
```

You want to know, for each student, how many negative outcomes were predicted? How many negative outcomes did they experience? Look at a 5x5 confusion matrix

Another way you could do this would be to compare the sum of the predictions - optimal_cutoffs to the number of negative outcomes. Look at a scatterplot.

```{r}
wgu_pa <- loadObject("wgu_preds_and_actuals")
ec_pa <- loadObject("ec_preds_and_actuals")

model_results <- loadObject("model_results_211028_1134")

load("ec_tidy.csv")
load("wgu_tidy.csv")
```

# WGU

```{r}

#This is just for for predictions from base models

wgu_pa_base <- wgu_pa %>%
  mutate("total_actual_positive_outcomes" = actual_retained + actual_ontime_term1 + actual_ontime_term2 + actual_creditratio_term1_bin + actual_creditratio_term2_bin) %>% mutate(
    "total_predicted_positive_outcomes" = (pred_wgu_base_Retained > 0.6177) +
      (pred_wgu_base_OnTime_Term1 > 0.5229) +
      (pred_wgu_base_OnTime_Term2 > 0.5242) +
      (pred_wgu_base_CreditRatio_Term1_bin > 0.4452) +
      (pred_wgu_base_CreditRatio_Term2_bin > 0.4454))

wgu_pa_daacs <- wgu_pa %>%
    mutate("total_actual_positive_outcomes" = actual_retained + actual_ontime_term1 + actual_ontime_term2 + actual_creditratio_term1_bin + actual_creditratio_term2_bin) %>% mutate(
    "total_predicted_positive_outcomes" = (pred_wgu_daacs_Retained > 0.5009) +
      (pred_wgu_daacs_OnTime_Term1 > 0.5640) +
      (pred_wgu_daacs_OnTime_Term2 > 0.4673) +
      (pred_wgu_daacs_CreditRatio_Term1_bin > 0.4894) +
      (pred_wgu_daacs_CreditRatio_Term2_bin > 0.4336))
```

```{r}
tmp <- wgu_pa_base %>%
  filter(!is.na(total_predicted_positive_outcomes))

nrow(tmp)

#actuals are rows, preds are cols
table(tmp$total_actual_positive_outcomes, tmp$total_predicted_positive_outcomes)
```
with the daacs model:
If we predict you will have no negative outcomes, then we're 49% accurate. (n = 53)
If we predict you will have 1 negative outcome, then there's a 63% chance you will have at least 1 negative outcome. (n = 228)
If we predict you will have 2 negative outcomes, then there's a 75% chance you will have at least 1 negative outcome. (n = 683)
If we predict you will have 3, then p(at least 1) is 87% (n = 142)
If we predict you will have 4, then p(at least 1) is 96% (n = 144)

with base:
If we predict you will have no negative outcomes, (n = 1)
If we predict you will have 1 negative outcome, then there's a 67% chance you will have at least 1 negative outcome. (n = 30)
If we predict you will have 2 negative outcomes, then there's a 76% chance you will have at least 1 negative outcome. (n = 1084)
If we predict you will have 3, then p(at least 1) is 86% (n = 56)
If we predict you will have 4, then p(at least 1) is 89% (n = 80)

We can make some claims about some students. But those claims are pretty hedged, and the number of students we can make strong claims about is small.

# EC

```{r}
#This is just for for predictions from base models

ec_pa_base <- ec_pa %>%
  mutate("total_actual_positive_outcomes" = actual_retained + actual_successterm1 + actual_successterm2 + actual_creditratio_term1_bin + actual_creditratio_term2_bin) %>% mutate(
    "total_predicted_positive_outcomes" = (pred_ec_base_Retained > 0.5070) +
      (pred_ec_base_SuccessTerm1 > 0.4737) +
      (pred_ec_base_SuccessTerm2 > 0.4960) +
      (pred_ec_base_CreditRatio_Term1_bin > 0.5840) +
      (pred_ec_base_CreditRatio_Term2_bin > 0.5255))

ec_pa_daacs <- ec_pa %>%
    mutate("total_actual_positive_outcomes" = actual_retained + actual_successterm1 + actual_successterm2 + actual_creditratio_term1_bin + actual_creditratio_term2_bin) %>% mutate(
    "total_predicted_positive_outcomes" = (pred_ec_daacs_Retained > 0.4181) +
      (pred_ec_daacs_SuccessTerm1 > 0.6299) +
      (pred_ec_daacs_SuccessTerm2 > 0.4561) +
      (pred_ec_daacs_CreditRatio_Term1_bin > 0.7362) +
      (pred_ec_daacs_CreditRatio_Term2_bin > 0.3982))
```

```{r}
tmp <- ec_pa_base %>%
  filter(!is.na(total_predicted_positive_outcomes))

nrow(tmp)

#actuals are rows, preds are cols
table(tmp$total_actual_positive_outcomes, tmp$total_predicted_positive_outcomes)
```
