---
title: "Binary Metrics"
author: "Daniel Moscoe"
date: "10/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


loadObject loads an object but allows you to set the name of the object in the environment.
binary_metrics is used to help populate the table models, which contains all the results from the SuperLearners.
preds_and_actuals were populated based on SuperLearner objects according to the code in preds_and_actuals.
```{r}
library(tidyverse)
library(InformationValue)
library(pROC)

loadObject <- function(fileName){
  load(fileName)
  get(ls()[ls() != "fileName"])
}

#Actuals, then predictions
binary_metrics <- function(df){
  optimal_cutoff <- InformationValue::optimalCutoff(actuals = df[1],
                                                    predictedScores = df[2],
                                                    optimiseFor = "misclasserror")
  
  naive_accuracy <- max(sum(df[1])/nrow(df), 1 - sum(df[1])/nrow(df))
  
  cm <- InformationValue::confusionMatrix(actuals = df[[1]],
                                          predictedScores = df[[2]],
                                          threshold = optimal_cutoff)
  auc <- pROC::auc(df[[1]], df[[2]])

  res <- data.frame("oc" = optimal_cutoff,
                    "auc" = auc,
                    "tp" = cm[2,2],
                    "fp" = cm[2,1],
                    "tn" = cm[1,1],
                    "fn" = cm[1,2],
                    "na" = naive_accuracy)
  return(res)
}
?read_csv
models <- read_csv("models.csv", col_types = "icccccddiiiid")
ec_preds_and_actuals <- loadObject("ec_preds_and_actuals")
wgu_preds_and_actuals <- loadObject("wgu_preds_and_actuals")
```

```{r}
#actuals is the column in a preds_and_actuals containing actual values.
#preds is the column in a preds_and_actuals containing predicted values.
#r is the row number in models to load with the results of binary_metrics.

loadModelsDF <- function(actuals, preds, r){
  df <- data.frame("actuals" = actuals,
                   "preds" = preds)
  df <- filter(df, !is.na(preds))
  to_load <- binary_metrics(df)
  models[r,c(7:13)] <<- to_load[1,]
}

```

This populates the table models with the binary metrics from each SuperLearner model.
```{r, message = FALSE, warning = FALSE}
loadModelsDF(wgu_preds_and_actuals$actual_ontime_term1, wgu_preds_and_actuals$pred_wgu_base_OnTime_Term1, 1)
loadModelsDF(wgu_preds_and_actuals$actual_ontime_term2, wgu_preds_and_actuals$pred_wgu_base_OnTime_Term2, 2)
loadModelsDF(wgu_preds_and_actuals$actual_retained, wgu_preds_and_actuals$pred_wgu_base_Retained, 3)
loadModelsDF(wgu_preds_and_actuals$actual_creditratio_term1_bin, wgu_preds_and_actuals$pred_wgu_base_CreditRatio_Term1_bin, 4)
loadModelsDF(wgu_preds_and_actuals$actual_creditratio_term2_bin, wgu_preds_and_actuals$pred_wgu_base_CreditRatio_Term2_bin, 5)
loadModelsDF(wgu_preds_and_actuals$actual_ontime_term1, wgu_preds_and_actuals$pred_wgu_daacs_OnTime_Term1, 8)
loadModelsDF(wgu_preds_and_actuals$actual_ontime_term2, wgu_preds_and_actuals$pred_wgu_daacs_OnTime_Term2, 9)
loadModelsDF(wgu_preds_and_actuals$actual_retained, wgu_preds_and_actuals$pred_wgu_daacs_Retained, 10)
loadModelsDF(wgu_preds_and_actuals$actual_creditratio_term1_bin, wgu_preds_and_actuals$pred_wgu_daacs_CreditRatio_Term1_bin, 11)
loadModelsDF(wgu_preds_and_actuals$actual_creditratio_term2_bin, wgu_preds_and_actuals$pred_wgu_daacs_CreditRatio_Term2_bin, 12)
loadModelsDF(ec_preds_and_actuals$actual_successterm1, ec_preds_and_actuals$pred_ec_base_SuccessTerm1, 15)
loadModelsDF(ec_preds_and_actuals$actual_successterm2, ec_preds_and_actuals$pred_ec_base_SuccessTerm2, 16)
loadModelsDF(ec_preds_and_actuals$actual_retained, ec_preds_and_actuals$pred_ec_base_Retained, 17)
loadModelsDF(ec_preds_and_actuals$actual_creditratio_term1_bin, ec_preds_and_actuals$pred_ec_base_CreditRatio_Term1_bin, 18)
loadModelsDF(ec_preds_and_actuals$actual_creditratio_term2_bin, ec_preds_and_actuals$pred_ec_base_CreditRatio_Term2_bin, 19)
loadModelsDF(ec_preds_and_actuals$actual_successterm1, ec_preds_and_actuals$pred_ec_daacs_SuccessTerm1, 22)
loadModelsDF(ec_preds_and_actuals$actual_successterm2, ec_preds_and_actuals$pred_ec_daacs_SuccessTerm2, 23)
loadModelsDF(ec_preds_and_actuals$actual_retained, ec_preds_and_actuals$pred_ec_daacs_Retained, 24)
loadModelsDF(ec_preds_and_actuals$actual_creditratio_term1_bin, ec_preds_and_actuals$pred_ec_daacs_CreditRatio_Term1_bin, 25)
loadModelsDF(ec_preds_and_actuals$actual_creditratio_term2_bin, ec_preds_and_actuals$pred_ec_daacs_CreditRatio_Term2_bin, 26)

```

This computes the accuracy of each superlearner model in the table models.
```{r}
models <- models %>%
  mutate("accuracy" = (true_positives + true_negatives) / (true_positives + true_negatives + false_positives + false_negatives))
```

```{r}
save(models, file = "model_results_211028_1134")
write_csv(models, file = "model_results_211028_1134.csv")
```

