---
title: "inference_211101_1558"
author: "Daniel Moscoe"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Are there significant differences in the number of actual positive outcomes for students across groups A and B?
Group A: students for whom the total number of predicted positive outcomes is highest
Group B: students for whom the total number of predicted positive outcomes is lowest

Compare mean total positive outcomes for groups A and B for each of:
WGU Base
WGU DAACS
EC Base
EC DAACS

```{r}
setwd("C:/Users/dmosc/OneDrive/Documents/academic/CUNY SPS/capstone/scripts_and_models_211026ff")

loadObject <- function(fileName){
  load(fileName)
  get(ls()[ls() != "fileName"])
}

wgu_pa <- loadObject("wgu_preds_and_actuals")
ec_pa <- loadObject("ec_preds_and_actuals_211102")

```

Compute the model prediction less the optimal cutoff value for each variable for the EC base and DAACS models
```{r}
ec_pa_excess_preds <- ec_pa %>%
  mutate("bin_pred_ec_base_CreditRatio_Term1" = pred_ec_base_CreditRatio_Term1_bin - 0.5840) %>%
  mutate("bin_pred_ec_base_CreditRatio_Term2" = pred_ec_base_CreditRatio_Term2_bin - 0.5255) %>%
  mutate("bin_pred_ec_base_SuccessTerm1" = pred_ec_base_SuccessTerm1 - 0.4737) %>%
  mutate("bin_pred_ec_base_Success_Term2" = pred_ec_base_SuccessTerm2 - 0.4960) %>%
  mutate("bin_pred_ec_base_Retained" = pred_ec_base_Retained - 0.5070) %>%
  mutate("bin_pred_ec_daacs_CreditRatio_Term1" = pred_ec_daacs_CreditRatio_Term1_bin - 0.7362) %>%
  mutate("bin_pred_ec_daacs_CreditRatio_Term2" = pred_ec_daacs_CreditRatio_Term2_bin - 0.3982) %>%
  mutate("bin_pred_ec_daacs_SuccessTerm1" = pred_ec_daacs_SuccessTerm1 - 0.6299) %>%
  mutate("bin_pred_ec_daacs_Success_Term2" = pred_ec_daacs_SuccessTerm2 - 0.4561) %>%
  mutate("bin_pred_ec_daacs_Retained" = pred_ec_daacs_Retained - 0.4181)
```

Compute the total number of actual positive outcomes and the total prediction - optimal cutoff
```{r}
ec_pa_excess_preds <- ec_pa_excess_preds %>%
  mutate("sum_actuals" = actual_retained + actual_successterm1 + actual_successterm2 + actual_creditratio_term1_bin + actual_creditratio_term2_bin) %>%
  mutate("sum_base_preds" = bin_pred_ec_base_CreditRatio_Term1 + bin_pred_ec_base_CreditRatio_Term2 + bin_pred_ec_base_SuccessTerm1 + bin_pred_ec_base_Success_Term2 + bin_pred_ec_base_Retained) %>%
  mutate("sum_daacs_preds" = bin_pred_ec_daacs_CreditRatio_Term1 + bin_pred_ec_daacs_CreditRatio_Term2 + bin_pred_ec_daacs_SuccessTerm1 + bin_pred_ec_daacs_Success_Term2 + bin_pred_ec_daacs_Retained)
```

```{r}
ec_pa_excess_preds <- arrange(ec_pa_excess_preds, sum_base_preds)
```

There are 507 rows with base predictions. They're sorted ascending. Compare mean sum_actuals for the first half of these rows with sum_actuals for second half of these rows.

```{r}
tmp_group_B <- ec_pa_excess_preds[c(1:253), "sum_actuals"]
tmp_group_A <- ec_pa_excess_preds[c(254:507), "sum_actuals"]
```

Is the mean sum_actuals for group A significantly greater than the mean sum_actuals for group B? Yes.
```{r}
t.test(tmp_group_A, tmp_group_B)
```
###211110###
```{r}
tmp <- data.frame("obs" = c(tmp_group_A, tmp_group_B), "group" = NA)
tmp[c(1:254),"group"] <- "A"
tmp[c(255:507),"group"] <- "B"

ggplot(tmp, aes(x = obs, fill = group)) +
  geom_density(alpha = 0.5)


hist(tmp_group_A)
hist(tmp_group_B)

```
###211110###

```{r}
ec_pa_excess_preds <- arrange(ec_pa_excess_preds, sum_daacs_preds)
```

There are 157 rows with DAACS predictions. They're sorted ascending. Compare mean sum_actuals for the first half of these rows with sum_actuals for the second half of these rows.

```{r}
tmp_group_B <- ec_pa_excess_preds[c(1:78), "sum_actuals"]
tmp_group_A <- ec_pa_excess_preds[c(79:157), "sum_actuals"]
```

Is the mean sum_actuals for group A significantly greater than the mean sum_actuals for group B? Yes, but the CI is much wider. We need more data here.
```{r}
t.test(tmp_group_A, tmp_group_B)
```

###211110###
```{r}
tmp <- data.frame("obs" = c(tmp_group_A, tmp_group_B), "group" = NA)
tmp[c(1:79),"group"] <- "A"
tmp[c(80:157),"group"] <- "B"

ggplot(tmp, aes(x = obs, fill = group)) +
  geom_density(alpha = 0.5)
```


Now we do the same thing for WGU.

```{r}
wgu_pa_excess_preds <- wgu_pa %>%
  mutate("bin_pred_wgu_base_CreditRatio_Term1" = pred_wgu_base_CreditRatio_Term1_bin - 0.4452) %>%
  mutate("bin_pred_wgu_base_CreditRatio_Term2" = pred_wgu_base_CreditRatio_Term2_bin - 0.4454) %>%
  mutate("bin_pred_wgu_base_OnTime_Term1" = pred_wgu_base_OnTime_Term1 - 0.5229) %>%
  mutate("bin_pred_wgu_base_OnTime_Term2" = pred_wgu_base_OnTime_Term2 - 0.5242) %>%
  mutate("bin_pred_wgu_base_Retained" = pred_wgu_base_Retained - 0.6177) %>%
  mutate("bin_pred_wgu_daacs_CreditRatio_Term1" = pred_wgu_daacs_CreditRatio_Term1_bin - 0.4894) %>%
  mutate("bin_pred_wgu_daacs_CreditRatio_Term2" = pred_wgu_daacs_CreditRatio_Term2_bin - 0.4336) %>%
  mutate("bin_pred_wgu_daacs_OnTime_Term1" = pred_wgu_daacs_OnTime_Term1 - 0.5640) %>%
  mutate("bin_pred_wgu_daacs_OnTime_Term2" = pred_wgu_daacs_OnTime_Term2 - 0.4673) %>%
  mutate("bin_pred_wgu_daacs_Retained" = pred_wgu_daacs_Retained - 0.5009)
```

Compute the total number of actual positive outcomes and the total prediction - optimal cutoff
```{r}
wgu_pa_excess_preds <- wgu_pa_excess_preds %>%
  mutate("sum_actuals" = actual_retained + actual_ontime_term1 + actual_ontime_term2 + actual_creditratio_term1_bin + actual_creditratio_term2_bin) %>%
  mutate("sum_base_preds" = bin_pred_wgu_base_CreditRatio_Term1 + bin_pred_wgu_base_CreditRatio_Term2 + bin_pred_wgu_base_OnTime_Term1 + bin_pred_wgu_base_OnTime_Term2 + bin_pred_wgu_base_Retained) %>%
  mutate("sum_daacs_preds" = bin_pred_wgu_daacs_CreditRatio_Term1 + bin_pred_wgu_daacs_CreditRatio_Term2 + bin_pred_wgu_daacs_OnTime_Term1 + bin_pred_wgu_daacs_OnTime_Term2 + bin_pred_wgu_daacs_Retained)
```

```{r}
wgu_pa_excess_preds <- arrange(wgu_pa_excess_preds, sum_base_preds)
```

There are 1252 rows with base predictions. They're sorted ascending. Compare mean sum_actuals for the first half of these rows with sum_actuals for second half of these rows.

```{r}
tmp_group_B <- wgu_pa_excess_preds[c(1:626), "sum_actuals"]
tmp_group_A <- wgu_pa_excess_preds[c(627:1252), "sum_actuals"]
```

Is the mean sum_actuals for group A significantly greater than the mean sum_actuals for group B? Yes.
```{r}
t.test(tmp_group_A, tmp_group_B)
```
###211110###
```{r}
tmp <- data.frame("obs" = c(tmp_group_A, tmp_group_B), "group" = NA)
tmp[c(1:626),"group"] <- "A"
tmp[c(627:1252),"group"] <- "B"

ggplot(tmp, aes(x = obs, fill = group)) +
  geom_density(alpha = 0.5)
```


```{r}
wgu_pa_excess_preds <- arrange(wgu_pa_excess_preds, sum_daacs_preds)
```

There are 1252 rows with DAACS predictions. They're sorted ascending. Compare mean sum_actuals for the first half of these rows with sum_actuals for the second half of these rows.

```{r}
tmp_group_B <- wgu_pa_excess_preds[c(1:626), "sum_actuals"]
tmp_group_A <- wgu_pa_excess_preds[c(627:1252), "sum_actuals"]
```

Is the mean sum_actuals for group A significantly greater than the mean sum_actuals for group B? Yes.
```{r}
t.test(tmp_group_A, tmp_group_B)
```
###211110###

```{r}
tmp <- data.frame("obs" = c(tmp_group_A, tmp_group_B), "group" = NA)
tmp[c(1:626),"group"] <- "A"
tmp[c(627:1252),"group"] <- "B"

ggplot(tmp, aes(x = obs, fill = group)) +
  geom_density(alpha = 0.5)
```

DAACS assessment results can help identify students most in need of academic intervention/enhanced advising/additional support.
I predicted five academic outcomes for each student using demographic data only, and then again using demographic and DAACS assessment data. Then I split students at each school into two groups. Group A had the highest number of predicted positive outcomes. That is, I predict these students will do the best. Group B had the lowest number of predicted positive outcomes. That is, I predict these students will do the worst.

At EC using base variables only, there was a difference in average number of positive outcomes of 0.50. Students in Group A had on average 3.47 positive outcomes. Students in Group B had on average 2.97 positive outcomes (p = 0.001).

At EC using base and DAACS variables, there was a difference in average number of positive outcomes of 0.62. Students in Group A had on average 3.71 positive outcomes. Students in Group B had on average 3.09 positive outcomes (p = 0.027).

At WGU using base variables only, there was a difference in average number of positive outcomes of 0.63. Students in Group A had on average 3.02 positive outcomes. Students in Group B had on average 2.39 positive outcomes (p < 0.001).

At WGU using base and DAACS variables, there was a difference in average number of positive outcomes of 1.00. Students in Group A had on average 3.23 positive outcomes. Students in Group B had on average 2.23 positive outcomes (p < 0.001).

DAACS assessments when combined with other data can help identify a cohort of students likely to attain fewer positive academic outcomes. This can help institutions target support resources to students who need them the most.

































